Validation of Visual Statistical Inference, with Application to Linear Models

### 2012-02-03 paper_jasa.tex
new update

%testing

@@

%The paper concludes by describing the plans for future experiments.

@@

%This section gives the background to the lineup protocol described by \citet{buja:2009}.
This section outlines the concepts of visual inference in comparison to the procedures of classical statistical inference. Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.

Let $\theta$ be a population parameter of interest, with $\theta \in \Theta$, the parameter space. 
Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$. 
% In hypothesis testing terminology, the parameter space $\Theta$ of a population parameter $\theta$, can be partitioned into $\Theta_0$ and $\Theta_0^c$. We test $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.

@@

For two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable visually. This should be reflected in the power of the test. We can assess power therefore both empirically based on experimental data and through theoretical considerations. Next, we will develop power theoretically and then %It is feasible to assess power theoretically and 
relate it to the empirical results.

@@

%\begin{table*}[hbtp] 
%	\caption{Test Statistics for Testing Hypothesis Related To The Model $Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_1 X_2+ \epsilon $ } 
%	\centering 
%		\begin{tabular}{m{3cm}m{2.5cm}m{2.6cm}m{5.5cm}} 
%			\hline\hline 
%				Null Hypothesis & Statistic Name & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
%			\hline 
%
%		\end{tabular} 
%	\label{tbl:stat_category} 
%\end{table*}

@@

Figure \ref{fig:power_expected} shows the power of UMP test and expected power of visual test obtained from equation \eqref{power_exp}. Notice that the expected power of visual test is almost as good as the power of UMP test. In fact, the only loss we experience stems from comparing the test statistic to only a discrete number $m-1$ samples of the null distribution. For infinitely many null plots we have the following theorem.

\begin{thm}
 Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_B)^{m-1}$ .
\end{thm}

@@

@@

%\begin{table}[hbtp] % test table
%\caption{Values of parameters considered for survey experiment} % title name of the table
%\centering
%  \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
%    \hline
%    \multicolumn{1}{|>{\columncolor{kugray5}}c|}{}&\multicolumn{4}{c|}{Rank}&\multirow{2}{*}{Total}\\
%    \arrayrulecolor{kugray5}
%    \arrayrulecolor{black}
%    \cline{2-5}
%    \multicolumn{1}{|>{\columncolor{kugray5}}c|}{}&A&B&C&Other&\\
%    \hline
%    type 2&8&14&5&2&29\\
%    \hline
%  \end{tabular}
%\end{table}

>>>>

@@

@@

This section outlines the concepts of visual inference in comparison to the procedures of classical statistical inference. Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.

Let $\theta$ be a population parameter of interest, with $\theta \in \Theta$, the parameter space. 
Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.

@@

For two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable visually. This should be reflected in the power of the test. We can assess power therefore both empirically based on experimental data and through theoretical considerations. Next, we will develop power theoretically and then relate it to the empirical results.

@@

@@

Figure \ref{fig:power_expected} shows the power of UMP test and expected power of visual test obtained from equation \eqref{power_exp}. Notice that the expected power of visual test is almost as good as the power of UMP test. In fact, the only loss we experience stems from comparing the test statistic to only a discrete number $m-1$ samples of the null distribution. Theorem \ref{thm:power} helps to compute the expected power of visual inference.

\begin{thm}\label{thm:power}
 Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_B)^{m-1}$ .
\end{thm}

@@

% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t

@@
